### 2011-07-26 paper_ecal/fitnessEvaluation.tex

As mentioned earlier, the metric for evaluating gaits was the Euclidian distance the robot
moved during a 12-second run. For the manual and parameterized gaits,
the fitness was this value. The HyperNEAT gaits stressed the motors
more than the other gaits, so to encourage gaits that did not tax the
motors we penalized gaits that caused the servos to stop
responding. When the servos stopped responding they could, in nearly
all cases, be restarted by cycling power, though over the course of
this study we did have to replace four servos that were damaged.
The penalty was to set the fitness to half of the distance
the robot actually traveled. We tested whether the servos were
responding after each gait by commanding them to specific positions
and checking whether they actually moved to those positions. This test
had the additional benefit of rewarding those gaits that did not flip
the robot into a position where it could not move its legs, which
HyperNEAT also did more than the other learning methods. Because the
fitness of HyperNEAT gaits were often halved, in results we compare
actual distance traveled in addition to fitness for the best gaits
produced by each class of gait-generating algorithms.

@@

In order to measure the start and end position in the same pose, and to ensure fair fitness evaluations with as little noise as possible, we
linearly interpolated the motion of the robot between the ready
position and the commanded gait, $g(t)$.  As shown in
\figref{linearInterp}, motion for the first second was interpolated
between the ready position and $g(t)$, for the next nine seconds the
robot moved exactly as specified by $g(t)$, and for the final two
seconds its motion was interpolated smoothly back to the ready
position.

@@

The only human intervention required during most learning trials was
to occasionally move the robot back into the viewable area of the Wii
remote whenever it left this window.  Initially this was a rare
occurrence, as the gaits did not typically produce motion as large as
the size of the window (roughly 175 x 120 cm).  However, as gaits
improved, particularly when using HyperNEAT, the robot began to walk
out of the measurement area a non-negligible fraction of the time.
Whenever it did so, we would discard the trial and repeat it until the
gait finished within the window. While this process guaranteed that we always obtained a measurement
for a given gait before proceeding, it also biased some measurements
downward.  Because the performance of the robot on a given gait varied
from trial to trial, a successful measurement was more likely to be
obtained when the gait happened to perform poorly.  This phenomenon
was negligible at first, but became more pronounced as gaits began traversing the entire
area.  HyperNEAT gaits were especially likely to require additional trials, meaning that the reported performance for HyperNEAT is worse than it would have been otherwise. For future studies we will use a $2\times2$ array of Wii
remotes to double the size of the measurement arena.

>>>>

As mentioned earlier, the metric for evaluating gaits was the Euclidian distance the robot
moved during a 12-second run on flat terrain. For the manual and parameterized gaits,
the fitness was this value. The HyperNEAT gaits stressed the motors
more than the other gaits, so to encourage gaits that did not tax the
motors we penalized gaits that caused the servos to stop
responding. When the servos stopped responding they could, in nearly
all cases, be restarted by cycling power, though over the course of
this study we did have to replace four servos that were damaged.
The penalty was to set the fitness to half of the distance
the robot actually traveled. We tested whether the servos were
responding after each gait by commanding them to specific positions
and checking whether they actually moved to those positions. This test
had the additional benefit of rewarding those gaits that did not flip
the robot into a position where it could not move its legs, which
HyperNEAT also did more than the other learning methods. Because the
fitness of HyperNEAT gaits were often halved, in results we compare
actual distance traveled in addition to fitness for the best gaits
produced by each class of gait-generating algorithms.

@@

In order to measure the start and end position in the same pose, and to ensure fair fitness evaluations with as little noise as possible, we
linearly interpolated the motion of the robot between the ready
position and the commanded gait, $g(t)$.  As shown in
%\figref{linearInterp}, motion for the first second was interpolated
%between the ready position and $g(t)$, for the next nine seconds the
%robot moved exactly as specified by $g(t)$, and for the final two
%seconds its motion was interpolated smoothly back to the ready
%position.
\figref{linearInterp}, the instantaneous robot limb configuration
during the first and last portions of the evaluation was an
interpolation between the initial ready position and $g(t)$;
during the rest of the evaluation, the robot followed the commanded
gait exactly.

@@

The only human intervention required during most learning trials was
to occasionally move the robot back into the viewable area of the Wii
remote whenever it left this window.  Initially this was a rare
occurrence, as the gaits did not typically produce motion as large as
the size of the window (roughly 175 x 120 cm).  However, as gaits
improved, particularly when using HyperNEAT, the robot began to walk
out of the measurement area a non-negligible fraction of the time.
Whenever it did so, we would discard the trial and repeat it until the
gait finished within the window. While this process guaranteed that we always obtained a measurement
for a given gait before proceeding, it also biased some measurements
downward.  Because the performance of the robot on a given gait varied
from trial to trial, a successful measurement was more likely to be
obtained when the gait happened to perform poorly.  This phenomenon
was negligible at first, but became more pronounced as gaits began traversing the entire
area.  HyperNEAT gaits were especially likely to require additional
trials, meaning that the reported performance for HyperNEAT is worse
than it would have been otherwise. Future studies could employ an
array of Wii remotes to increase the size of the measurement arena.

### 2011-07-26 poster_gecco/poster.tex

    The quadruped robot was assembled from off-the-shelf
    components and parts printed on the Objet Connex 500 3-D Printing
    System.
    % It weighs 1.88 kg with the on-board computer and measures 38
    % centimeters from leg to opposite leg in the crouch position
    % depicted above.
    The robot is actuated by 9 AX-12+ Dynamixel servos: 8 for the legs and one at the center ``hip'' joint.
    %This final unique servo allows
    %the two halves of the robot to rotate with respect to each other.

@@

\pfigg{wiiMote_crop.jpg}{robot_led_crop.jpg}{1}{A Nintendo Wii
  remote tracked an IR LED on the robot, transmitting the (x,y)
  location via bluetooth.}

@@

The fitness was the Euclidian distance the robot moved during a
12-second run on flat terrain.  To encourage gaits that did not tax
the motors, we penalized gaits that caused the servos to stop
responding by a factor of 1/2.

@@

The robot was always measured while in the \emph{ready} position
to ensure fair fitness measurements. As
shown above the instantaneous robot limb configuration during the
first and last portions of the evaluation was an interpolation between
the initial ready position and $g(t)$.

@@

    \pfig{neat_110115_211410_00000_002_filt_zoom}{1}{Joint angles for
      a high-performance HyperNEAT gait.}

@@

\noindent A total of 1217 hardware fitness evaluations were performed to learn
parameterized gaits in three separate runs (200 random, 234 uniform, 284 Gaussian, 174
gradient, 172 simplex, 153 linear regression). The SVM-modeled method
was never successful.

@@

Evaluation of HyperNEAT gaits was very noisy, which made learning
difficult. We tested one gait 11 times and found performance of 26
$\pm$ 13 body lengths/minute, with a max of 38 and a min of 3.

@@

\textbf{Overall}
\vspace{.5em}

>>>>

    The quadruped robot was assembled from off-the-shelf
    components and parts printed on the Objet Connex 500 3-D Printing
    System.
    % It weighs 1.88 kg with the on-board computer and measures 38
    % centimeters from leg to opposite leg in the crouch position
    % depicted above.
    The robot is actuated by nine AX-12+ Dynamixel servos: eight for the legs and one at the center ``hip'' joint.
    %This final unique servo allows
    %the two halves of the robot to rotate with respect to each other.

@@

\pfigg{wiiMote_crop.jpg}{robot_led_crop.jpg}{1}{\\A Nintendo Wii
  remote tracked an IR LED on the robot, transmitting the (x,y)
  location via bluetooth.}

@@

    Gait fitness was the Euclidian distance the robot moved during a
    12-second run on flat terrain.  To encourage gaits that did not
    tax the motors, we penalized gaits that caused the servos to stop
    responding by a factor of 1/2.

@@

    The robot was always measured while in the \emph{ready} position
    to ensure fair fitness measurements. As shown above, the
    instantaneous robot limb configuration during the first and last
    portions of the evaluation was an interpolation between the
    initial ready position and $g(t)$.

@@

    \pfig{neat_110115_211410_00000_002_filt_zoom}{1}{Joint angles over time for
      a high-performance HyperNEAT gait.}

@@

\noindent A total of 1217 hardware fitness evaluations were performed to learn
parameterized gaits in three separate runs (200 random, 234 uniform, 284 Gaussian, 174
gradient, 172 simplex, 153 linear regression). The SVM-based self-modeling method
was never successful.

@@

Evaluation of HyperNEAT gaits was very noisy, which made learning
difficult. For example, we tested one gait 11 times and found an average performance of 26
 body lengths/minute, $\pm$13 std. dev., with a max of 38 and a min of 3.

@@

\textbf{Summary of Results}
\vspace{.5em}
